A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-15 tex/codatadp.tex
Refactor proof for stronger lemma.

If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$, 
then $\M( t )$ is bisimilar with $\nf{(\mu x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
which by the induction hypothesis is bisimilar with $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
Since \rn{Clash} does not apply to $\Fc$,
then $\Val^\ast \ec{t}$ must be a term $t_w$ of the form $\mu v_{\ec{t}}. \const{C}( w_1, \ldots, w_n )$,
that is, an application of the same constructor.
For each $i = 1, \ldots, n$, let $u_i'$ be the $i^{th}$ argument of $t_u$.
By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$.
Say that $\muvar( w_i ) = v_{\ec{y_i}}$.
Due to our construction of $\Val^\ast$ and since \rn{Decomp} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{y_i}$.
By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{ y_i } = t_w\interp{ w_i }$,
and thus $t_u\interp{u_i'} \vsim t_w\interp{ w_i }$ for each $i = 1, \ldots, n$.
By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
Since $\vsim$ is a transitive relation and since $\M( t ) \vsim t_u \vsim t_w = \Val^\ast \ec{t}$,
we have $\M( t ) \vsim \Val^\ast \ec{t}$.

If $t$ is a selector term of the form $s^j_k( u )$,
since \rn{Split} does not apply, 
$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{n}_i( u ) )$ for some $i$.
By our construction of $\Val^\ast$ and since \rn{Decomp} and \rn{Clash} do not apply to $\Fc$, 
$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
where $\muvar( u_1 ) = v_{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = v_{\ec{s^n_i( u )}}$,
and $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
which by the induction hypothesis is bisimilar with $\Val^\ast \ec{u}$.
If $k = i$, then $\M( t )$ is bisimilar with $t_w\interp{w_j}$,
which since $t_w \vsim t_u$ by Lemma~\ref{lem:mu-cong} is bisimilar with $t_u\interp{u_j}$,
which by Lemma~\ref{lem:model-completion}.2 is equal to $\Val^\ast \ec{ s^j_k( u )}$.
Thus, $\M( t ) \vsim \Val^\ast \ec{ s^j_k( u )} = \Val^\ast \ec{t}$.
If $k \neq i$, then we interpret $s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is bisimilar with $\Val^\ast \ec{ t }$.
Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3, 
pairs of wrongly-applied selectors are assigned the same value if and only if they reside in the same equivalence class,
and thus this definition is well-defined.

>>>>

If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$, 
then $\M( t )$ is bisimilar with $\nf{(\mu x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
which by the induction hypothesis is bisimilar with $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{t}$ is a term
%Since \rn{Clash} does not apply to $\Fc$, then $\Val^\ast \ec{t}$ must be a term
$t_w$ of the form $\mu v_{\ec{t}}. \const{C}( w_1, \ldots, w_n )$
where $t_w\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
%that is, an application of the same constructor.
For each $i = 1, \ldots, n$, let $u_i'$ be the $i^{th}$ argument of $t_u$.
By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$.
%Say that $\muvar( w_i ) = v_{\ec{y_i}}$.
%Due to our construction of $\Val^\ast$ and since \rn{Decomp} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
%and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{y_i}$.
%By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{ y_i } = t_w\interp{ w_i }$,
and thus $t_u\interp{u_i'} \vsim t_w\interp{ w_i }$. % for each $i = 1, \ldots, n$.
By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
Since $\vsim$ is a transitive relation and since $\M( t ) \vsim t_u \vsim t_w = \Val^\ast \ec{t}$,
we have $\M( t ) \vsim \Val^\ast \ec{t}$.

If $t$ is a selector term of the form $s^j_k( u )$,
since \rn{Split} does not apply, 
$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{n}_i( u ) )$ for some $i$.
By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{u}$ is a term $t_w$ of the form
$\mu v_{\ec{u}}. C_i( w_1, \ldots, w_n )$,
where $t_w \interp{ w_j } = \Val^\ast \ec{ s^j_k( u )}$. 
%By our construction of $\Val^\ast$ and since \rn{Decomp} and \rn{Clash} do not apply to $\Fc$, 
%$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
%where $\muvar( u_1 ) = v_{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = v_{\ec{s^n_i( u )}}$,
%and $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
%which by the induction hypothesis is bisimilar with $\Val^\ast \ec{u}$.
If $k = i$, then $\M( t )$ is bisimilar with $t_w\interp{w_j} = \Val^\ast \ec{ s^j_k( u )}$,
which is $\Val^\ast \ec{t}$.
%which since $t_w \vsim t_u$ by Lemma~\ref{lem:mu-cong} is bisimilar with $t_u\interp{u_j}$,
%which by Lemma~\ref{lem:model-completion}.2 is equal to $\Val^\ast \ec{ s^j_k( u )}$.
%Thus, $\M( t ) \vsim \Val^\ast \ec{ s^j_k( u )} 
If $k \neq i$, then we interpret $s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is bisimilar with $\Val^\ast \ec{ t }$.
Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3, 
pairs of wrongly-applied selectors are assigned the same value if and only if they reside in the same equivalence class,
and thus this definition is well-defined.
